---
title: "ETC4500/5450 Advanced R Programming"
page-layout: full
execute:
  cache: false
  freeze: false
---

R is widely used as a tool for data analysis and one of the most popular programming languages. This unit delves into R from the programming aspect instead of using it as a data analysis tool. You will learn a variety of programming paradigms and delve into the foundations of R as a programming language.

::: {.action-grid}
<a class="action-card" href="https://learning.monash.edu/course/view.php?id=40292&section=1">
  <span class="action-card__kicker">Start here</span>
  <strong>Moodle</strong>
  <span>Lecture material, submissions, announcements, and unit administration.</span>
</a>
<a class="action-card" href="https://learning.monash.edu/mod/lti/view.php?id=5219874">
  <span class="action-card__kicker">Watch</span>
  <strong>Recordings</strong>
  <span>Pre-class videos and workshop recordings in one place.</span>
</a>
<a class="action-card" href="https://edstem.org/au/courses/31503/discussion/">
  <span class="action-card__kicker">Ask questions</span>
  <strong>Discussion</strong>
  <span>Course discussion, clarifications, and workshop support on Ed.</span>
</a>
<a class="action-card" href="#weekly-schedule">
  <span class="action-card__kicker">Plan ahead</span>
  <strong>Weekly schedule</strong>
  <span>Topics, week pages, and upcoming assessment milestones.</span>
</a>
:::

## Course at a glance

::: {.metric-grid}
<div class="metric-card">
  <span class="metric-card__label">Workshop</span>
  <strong>Thursday 9:00am to 11:00am</strong>
  <span>LTB Room 121</span>
</div>
<div class="metric-card">
  <span class="metric-card__label">Format</span>
  <strong>Pre-class videos + workshop</strong>
  <span>Arrive having watched the material for the week.</span>
</div>
<div class="metric-card">
  <span class="metric-card__label">Focus</span>
  <strong>Programming in R</strong>
  <span>Functional programming, packages, OO systems, metaprogramming, and performance.</span>
</div>
<div class="metric-card">
  <span class="metric-card__label">Assessment</span>
  <strong>3 assignments + discussion</strong>
  <span>Three major tasks worth 90% plus 10% discussion participation.</span>
</div>
:::

## Learning outcomes

::: {.outcome-grid}
<div class="outcome-card">
  <span class="outcome-card__number">01</span>
  <p>Build a solid understanding of the foundations of R as a programming language.</p>
</div>
<div class="outcome-card">
  <span class="outcome-card__number">02</span>
  <p>Work confidently across multiple paradigms, including functional and object-oriented programming.</p>
</div>
<div class="outcome-card">
  <span class="outcome-card__number">03</span>
  <p>Use R's functional and object-oriented systems to design reusable, well-structured code.</p>
</div>
<div class="outcome-card">
  <span class="outcome-card__number">04</span>
  <p>Apply metaprogramming ideas to evaluate, generate, and extend code programmatically.</p>
</div>
:::

## Teaching team and references

::: {.split-grid}
<div class="resource-card">
  <span class="resource-card__kicker">Teaching team</span>
  <h3>Who is teaching this unit</h3>
  <ul>
    <li><a href="https://robjhyndman.com"><strong>Rob J Hyndman</strong></a>. Chief Examiner. Email: <a href="mailto:Rob.Hyndman@monash.edu">Rob.Hyndman@monash.edu</a></li>
    <li><a href="https://krisanata.github.io/personal-website/"><strong>Krisanat Anukarnsakulchularp</strong></a>, Tutor</li>
    <li><a href="https://github.com/donotdespair"><strong>Tomasz Wozniak</strong></a>, Week 11 guest instructor</li>
  </ul>
</div>
<div class="resource-card">
  <span class="resource-card__kicker">Core references</span>
  <h3>Readings worth keeping open</h3>
  <ul>
    <li><a href="https://adv-r.hadley.nz/"><em>Advanced R</em>, 2e</a> by Hadley Wickham</li>
    <li><a href="https://advanced-r-solutions.rbind.io/"><em>Advanced R Solutions</em></a> by Malte Grosser, Henning Bumann and Hadley Wickham</li>
    <li><a href="http://r-pkgs.org/"><em>R packages</em>, 2e</a> by Hadley Wickham and Jenny Bryan</li>
  </ul>
</div>
:::

```{r}
#| label: load_packages
#| include: false
#| message: false
#| warning: false
#| echo: false
#| cache: false
library(tidyverse)
options(knitr.kable.NA = '')
source(here::here("course_info.R"))
```

## Weekly schedule

Each week combines pre-class material with a live workshop. Use the week pages for readings, activities, slides, and assignment reminders.

```{r}
#| label: schedule2
#| message: false
#| warning: false
#| echo: false
#| output: asis
#| cache: false
schedule |>
  mutate(
    week_num = Week,
    Week = if_else(
      is.na(week_num),
      "<span class='week-pill week-pill--break'>Break</span>",
      glue::glue("<span class='week-pill'>Week {week_num}</span>")
    ),
    Date = format(Date + 3, "%d %b"),
    Topic = case_when(
      is.na(week_num) ~ "<span class='muted'>Mid-semester break</span>",
      TRUE ~ glue::glue("<a href='./week{week_num}/index.html'>{Topic}</a>")
    ),
    Assessments = case_when(
      !is.na(Assignment) ~ glue::glue("<a class='mini-link' href='{File}'>{Assignment}</a>"),
      TRUE ~ "<span class='muted'>&mdash;</span>"
    )
  ) |>
  select(Week, Date, Topic, Assessments) |>
  knitr::kable(
    format = "html",
    escape = FALSE,
    table.attr = "class='table course-table'"
  )
```

## Assessment overview

::: {.assessment-grid}
<a class="assessment-card" href="assignments/A1.html">
  <span class="assessment-card__kicker">Assessment 1</span>
  <strong>Assignment 1</strong>
  <span>30% · Group package project</span>
</a>
<a class="assessment-card" href="assignments/A2.html">
  <span class="assessment-card__kicker">Assessment 2</span>
  <strong>Assignment 2</strong>
  <span>30% · Debugging and code-quality task</span>
</a>
<a class="assessment-card" href="assignments/A3.html">
  <span class="assessment-card__kicker">Assessment 3</span>
  <strong>Assignment 3</strong>
  <span>30% · Extension project with advanced features</span>
</a>
<a class="assessment-card" href="assignments/discussion.html">
  <span class="assessment-card__kicker">Engagement</span>
  <strong>Discussion</strong>
  <span>10% · Contribution through class discussion</span>
</a>
:::
