---
title: "Week 11: Rewriting R code in C++"
summary: "Rewrite performance-critical code in C++ using Rcpp and related tooling."
prepare: "Make sure your compiler toolchain is installed before class so you can complete the exercises."
resource_label: "Rcpp FAQ"
resource_url: "https://cran.r-project.org/package=Rcpp/vignettes/Rcpp-FAQ.pdf"
focus:
  - "Rcpp for integrating C++ with R"
  - "Compiled code inside packages"
  - "Hands-on coding exercises"
  - "Working with RcppArmadillo"
---

Guest Lecturer: [Tomasz Wozniak](https://github.com/donotdespair)

```{r}
week <- 11
source(here::here("course_info.R"))
```

```{r}
#| output: asis
show_week_overview(week)
```

## Main reference

* [25. *Advanced R*](https://adv-r.hadley.nz/rcpp.html)

## What you will learn this week

* **C++** for applications in **R** with [Tomasz Woźniak](https://github.com/donotdespair)
* The first steps with **Rcpp**
* Some stats with **RcppArmadillo**
* Hands-on coding with four exercises
* An **R** package with compiled code

## Preparation to the session

You must have a **cpp** compiler installed on your computer. Follow the instructions from [Section 1.3. by Eddelbuettel & François (2023)](https://cran.r-project.org/package=Rcpp/vignettes/Rcpp-FAQ.pdf). In short: 

* **Windows:** install [RTools](https://CRAN.R-project.org/bin/windows/Rtools/)
* **macOS:** install [Xcode Command Line Tools](https://www.freecodecamp.org/news/install-xcode-command-line-tools/)
* **Linux:** install the standard development packages


## Other references used during the session

* **Armadillo** library [documentation](https://arma.sourceforge.net/docs.html)
* Tsuda, M., [*Rcpp for everyone*](https://teuder.github.io/rcpp4everyone_en/)

## Scripts used during the session

* Simple list creation [`nicetry.cpp`](nicetry.cpp)
* Simple linear regression example [`nicelr.cpp`](nicelr.cpp)
* Simple loop application [`nicelist.cpp`](nicelist.cpp)
* Sampling from inverted gamma 2 distribution [`nicerig2.cpp`](nicerig2.cpp)
* Step-by-step package creation [`nicepackage.R`](https://github.com/donotdespair/15steps2nicepackage)

```{r}
#| output: asis
show_slides(week)
show_assignments(week)
```
