---
title: "Week 2: Foundations of R Programming"
summary: "Work through subsetting, control flow, functions, environments, and conditions in base R."
prepare: "Watch the subsetting videos and finish the `linear()` warm-up exercise before class."
resource_label: "Advanced R foundations"
resource_url: "https://adv-r.hadley.nz/foundations-intro.html"
focus:
  - "Subsetting vectors, lists, and matrices"
  - "Control flow and function writing"
  - "Environments and conditions"
  - "Writing functions with sensible defaults and errors"
---

```{r}
week <- 2
source(here::here("course_info.R"))
```

```{r}
#| output: asis
show_week_overview(week)
```

## Main reference

 * [4-8, *Advanced R*](https://adv-r.hadley.nz/foundations-intro.html)


## What you will learn this week

* Subsetting
* Control flow
* Functions
* Environments
* Conditions

## Pre-class activities

* Watch the following series of short videos on **subsetting in R**:

  * [Subsetting R objects: Basics](https://www.youtube.com/watch?v=Plcy3qM8wEE)
  * [Subsetting R objects: Lists](https://www.youtube.com/watch?v=7j6C1w9x9hw)
  * [Subsetting R objects: Matrices](https://www.youtube.com/watch?v=83NSymr6pPw)
  * [Subsetting R objects: Partial matching](https://www.youtube.com/watch?v=Kb4lXR6Z-Hc)
  * [Subsetting R objects: Removing missing values](https://www.youtube.com/watch?v=PYpIVFS6xwU)

* Write a function `linear`, that takes arguments `x`, `a` and `b`, and returns `a + b * x`. Give default values `a=0` and `b=1`, and make sure your function has sensible error messages when the inputs are not as expected. Your function should work when `x` is a numerical scalar, vector, or matrix.


```{r}
#| output: asis
show_slides(week)
```

[Solutions to class exercises](/week2/exercises2.R)

```{r}
#| output: asis
show_assignments(week)
```
